Relationships in the extremes and their influence on competition and coexistence

### 2021-07-08 paper/Paper.Rnw
added some  intro

>>>>

The maintenance of biodiversity in competitive ecosystems is a central idea in ecology. Among others, the modern theory of coexistence seeks to assess and quantify the mechanisms that result in coexisting species and thus the maintenance of biodiversity. 
Many are familiar with the principle of competitive exclusion - an idea which postulates if species within the same trophic level and similar niches exist, one will eventually outcompete the other (Hardin 1960). Modern coexistence theory provides a quantitative breakdown of contributing mechanisms as criteria for coexistence and competitive exclusion (Chesson 1994).

Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity. Recent work has shown that tail associations, or associations between the tail ends of data, and specifically the asymmetry between tail associations have important considerations for ecological phenomena; although, to our knowledge, the consequences/implications of asymmetric tail associations for competition and coexistence have not yet been investigated. 
Correlation measures such as Pearson or Spearman correlation disguise the asymmetry of tail associations, and datasets with different tail associations can have the same correlation coefficient (see Figure). Asymmetry in tail associations describes tighter correlations in regions of a dataset compared to other regions of the dataset. A simple example of asymmetric tail associations is a dataset which is more correlated in the lower (upper) half than the upper (lower) half, described as left (right) tail associated. Current research has shown that ATAs have important considerations for extinction risks, “skewness of spatial averages”, Taylor’s Law, Moran Effect/ spatial synchrony (Gosh et al 2020 Ecosphere, Gosh et al 2020 And. EcoRes,). Despite the existence of tail associations within ecological data, to our knowledge, the implications of tail associations for competition and coexistence has not yet been examined; doing so is the main goal of this paper.

Asymmetries in tail associations have the potential to influence competition and coexistence via one particular mechanism, namely, storage effects. Covariation between environmental fluctuations and competitive pressures allows for storage effect to be a contributing stabalizing mechanism to coexistence. Because the contributive value of storage effects relies on the association of competition and environment (Chesson 1994, Barbaras ..), it is reasonable to suspect the contribution of storage effects to coexistence can be altered by an association between competition and environment that is asymmetric. While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coesustence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms on coexistence (due to the fact that storage effect is the only mechanism that depends of the variance between two drivers).

While there are multiple studies on the influence of extremes (due to climate change), the influence of relationships in their extremes (also due to climate change), has not yet been explored. The lottery model within the framework of coexistence theory provides a suitable opportunity to explore extreme relationships and the consequences to coexistence.
